---
title: "Assignment"
author: "Alejandro Mendez"
date: "2/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Libraries

```{r}
library(tidyr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(scales)
library(timeDate)
library(ggpubr)

```


# Loading and preprocessing the data

**Data get from de link provided**
```{r}
URL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
filename <- paste("./data", "/", "data.zip", sep = "")
dir.create("data")
download.file(url = URL, destfile = filename)
unzip(zipfile = filename, exdir = "./data")


```

**Load the data**
```{r}
data <- read.csv("data/activity.csv")
data$date = as.Date(data$date, "%Y-%m-%d")
colnames(data)[2] <- "fechas"
str(data$fechas)
head(data)
```

# What is mean total number of steps taken per day?




**Make the histogram**
```{r}
g <- data %>% mutate(date_col = date(fechas)) %>%
  group_by(date_col) %>%
  summarize(value = sum(steps, na.rm = TRUE)) %>%
  ggplot(aes(date_col, value)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(title = "Total number of steps taken each day",
       x = "Date", y = "Steps") +
  scale_x_date(breaks = breaks_pretty(30))+ 
  theme(axis.text.x = element_text(angle = 90))
g
```


**Calculate the mean and median**
```{r}
mm <- data %>% mutate(date_col = date(fechas)) %>%
  group_by(date_col) %>%
  summarize(median_steps = median(steps, na.rm = TRUE),
            mean_steps =  mean(steps, na.rm = TRUE)) 
mm

```

# What is the average daily activity pattern?

**Make the plot**
```{r}
ts <- data %>% group_by(interval) %>%
  summarize(mean_steps =  mean(steps, na.rm = TRUE)) %>%
  subset(!is.na(mean_steps)) %>%
  ggplot(aes(interval, mean_steps)) +
  geom_line( color = "steelblue") +
  theme_minimal() +
  labs(title = "Mean of steps taken each interval",
       x = "Interval", y = "Steps") +
  theme(axis.text.x = element_text(angle = 90))+
  scale_x_continuous(breaks = pretty(data$interval, n = 20)) 
  

ts
```


**Get the interval who contain the maximum number of steps**
```{r}
mint <- data %>% group_by(interval) %>%
  summarize(mean_steps =  mean(steps, na.rm = TRUE)) %>%
  top_n(1)
mint

```


# Imputing missing values

**We have  missing data**

```{r}
sum(is.na(data))
```


**First, impute the NA values using the mean by interval group. Make nona dataframe with imputed values.**

```{r}
nona <- data
nona <- nona %>% group_by(interval) %>%
  mutate_all(funs(ifelse(is.na(.), mean(., na.rm = TRUE), .)))
nona$fechas <- data$fechas
sum(is.na(nona))
```

**Then, compare the histograms**
```{r}
gnona <- nona %>% mutate(date_col = date(fechas)) %>%
  group_by(date_col) %>% summarize(value= sum(steps)) %>%
  ggplot(aes(date_col, value)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(title = "Total number of steps taken each day",
       x = "Date", y = "Steps") +
  scale_x_date(breaks = breaks_pretty(30))+ 
  theme(axis.text.x = element_text(angle = 90))


ggarrange(g,gnona)
```

**And obtain mean and median again.**
```{r}
mmnona <- nona %>% mutate(date_col = date(fechas)) %>%
  group_by(date_col) %>%
  summarize(median_steps = median(steps, na.rm = TRUE),
            mean_steps =  mean(steps, na.rm = TRUE)) 
mmnona

```

**Before we didnt get the data to calculate mean and median on certain days. Now we replace this NA values with the median.**

# Are there differences in activity patterns between weekdays and weekends?

```{r}
datweek = data
datweek$fecha = isWeekday(data$fechas, wday = 1:5)
datweek$fecha = ifelse(datweek$fecha == TRUE, "Weekday", "Weekend")

pp <- datweek %>% group_by(fecha, interval) %>%
  summarize(mean_steps = mean(steps, na.rm = TRUE)) %>%
  ggplot(aes(interval, mean_steps )) +
  geom_line( color = "steelblue") +
  facet_wrap(~fecha, nrow = 2) +
  labs(title = "Average number of steps taken per 5-minute interval across weekdays and weekends")

pp
```

